\mbox{Specifying Imperative Higher Order Programs}

### 2009-09-30 papers/esop2010/main.tex
go

\newcommand{\wpre}{ {\bf wp}}
\newcommand{\sep}{ ~|~ }
\newcommand{\letml}{{\bf let}}
\newcommand{\inml}{{\bf in}}
\newcommand{\ifml}{{\bf if}}
\newcommand{\thenml}{{\bf then}}
\newcommand{\elseml}{{\bf else}}
\newcommand{\refml}{{\bf ref}}
\newcommand{\propml}{{\bf prop}}
\newcommand{\boolml}{{\bf bool}}
\newcommand{\recml}{{\bf rec}}
\newcommand{\efft}[1]{ \langle #1 \rangle }
\newcommand{\lift}[1]{ \langle #1 \rangle }
\newcommand{\alist}[1]{ \overline{#1} }
\newcommand{\Trueml}{{\bf True}}
\newcommand{\Falseml}{{\bf False}}
\newcommand{\trueml}{{\bf true}}
\newcommand{\falseml}{{\bf false}}
\newcommand{\coq}{Coq}

@@

@@

@@

\begin{figure}[htpb]
  \begin{eqnarray*}
    \ceil{α} &=& α\\
    \ceil{τ\ref_ρ} &=& \ceil{τ}\ref_ρ\\
    \ceil{τ -> τ'} &=& \ceil{τ} -> \ceil{τ'}\\
    \ceil{τ \times τ'} &=& \ceil{τ} \times-> \ceil{τ'}\\
    \ceil{\efft{ε}} &=& \efft{ε}\\
    \ceil{τ ->^ε τ'} &=& (\ceil{τ} -> \efft{ε} -> \propml)\times (\ceil{τ} ->
    \efft{ε} -> \efft{ε} -> \ceil{τ'} -> \propml)()
  \end{eqnarray*}
  \caption{Lifting types}
  \label{fig:typelift}
\end{figure}

Again in order to simplify future definitions, the definition typing relation
differs somewhat from the standard way (see Fig.\ref{fig:typing}). First of
all, we present the language in {\em A-normal form} (ANF), where all values
obtained by intermediate computations are named. However, our definition of
values is not syntactical: our ``semantic'' values are objects which are
either syntactic values (abstractions and variables) or are obtained by {\em
pure application}, the application which involves the pure arrow $τ->τ$.
Furthermore, the system is a type and effect system. There are in fact two
typing relations, the first one is of the form $Γ|-_v  t : τ$. In this case,
$t$ is a semantic value of type $τ$ and cannot have any effect. It can be seen
that the set of terms described by $|-_v$ is the precisely the one obtained by
combining syntactic values and pure application in a well-typed way.

@@

unusual : 
* ANF 
* values defined by types
* two different types of arrows
* distinction bool and prop
* notion of substitution - correspondence χ κ

missing : 
* value injection
* subeffecting
* logic typing

@@

@@

>>>>

\newcommand{\wpre}{ {\bf wp}}
\newcommand{\sep}{ ~|~ }
\newcommand{\letml}{{\bf let}}
\newcommand{\inml}{{\bf in}}
\newcommand{\ifml}{{\bf if}}
\newcommand{\thenml}{{\bf then}}
\newcommand{\elseml}{{\bf else}}
\newcommand{\refml}{{\bf ref}}
\newcommand{\propml}{{\bf prop}}
\newcommand{\boolml}{{\bf bool}}
\newcommand{\recml}{{\bf rec}}
\newcommand{\efft}[1]{ \langle #1 \rangle }
\newcommand{\lift}[1]{ \langle #1 \rangle }
\newcommand{\alist}[1]{ \overline{#1} }
\newcommand{\Trueml}{{\bf True}}
\newcommand{\Falseml}{{\bf False}}
\newcommand{\trueml}{{\bf true}}
\newcommand{\falseml}{{\bf false}}
\newcommand{\unitml}{{\bf unit}}
\newcommand{\coq}{Coq}

@@

also one or two examples in the described language

characteristics : 
\begin{itemize}
  \item Hoare logic
  \item logic and program share values  
  \item Type and effect system
  \item Region typing to track references
  \item the wp needs absence of aliasing to produce good POs
\end{itemize}

@@

technical presentation:
\begin{itemize}
  \item ANF
  \item values defined by typing 
  \item  two different types of arrows
  \item  distinction bool and prop
  \item  notion of substitution - correspondence χ κ
  \item semantics is standard (goes to appendix )
\end{itemize}

@@

\begin{figure}[htpb]
  \begin{eqnarray*}
    \ceil{α} &=& α\\
    \ceil{τ~\refml_ρ} &=& \ceil{τ}\refml_ρ\\
    \ceil{τ -> τ'} &=& \ceil{τ} -> \ceil{τ'}\\
    \ceil{τ \times τ'} &=& \ceil{τ} \times-> \ceil{τ'}\\
    \ceil{\efft{ε}} &=& \efft{ε}\\
    \ceil{τ ->^ε τ'} &=& (\ceil{τ} -> \efft{ε} -> \propml)\times (\ceil{τ} ->
    \efft{ε} -> \efft{ε} -> \ceil{τ'} -> \propml)()
  \end{eqnarray*}
  \caption{Lifting types}
  \label{fig:typelift}
\end{figure}

Again in order to simplify future definitions, the definition typing relation
differs somewhat from the standard way (see Fig.~\ref{fig:typing}). First of
all, we present the language in {\em A-normal form} (ANF), where all values
obtained by intermediate computations are named. However, our definition of
values is not syntactical: our ``semantic'' values are objects which are
either syntactic values (abstractions and variables) or are obtained by {\em
pure application}, the application which involves the pure arrow $τ->τ$.
Furthermore, the system is a type and effect system. There are in fact two
typing relations, the first one is of the form $Γ|-_v  t : τ$. In this case,
$t$ is a semantic value of type $τ$ and cannot have any effect. It can be seen
that the set of terms described by $|-_v$ is the precisely the one obtained by
combining syntactic values and pure application in a well-typed way.

@@

unusual : 
missing : 
* value injection
* subeffecting
* logic typing

@@

\subsection{Sharing between logic and programs}

To obtain sharing between
\subsection{References}

We haven't mentioned references yet, but all we need to support them is to add
the three usual functions $ref, :=, !$ for reference creation, assignment and
lookup, respectively, to the environment. In our system, these functions have
the following types:
\begin{eqnarray*}
  ref &:& ∀αρ.α ->^ρ α~\refml_ρ\\
  := &:& ∀αρ.α~\refml_ρ ->^\emptyset α ->^ρ \unitml\\
  ! &:& ∀αρ. α~\refml_ρ ->^ρ α
\end{eqnarray*}

\subsection{Alias restriction}

\begin{figure}[htpb]
  \begin{mathpar}
    {\inferrule*[Left=Var]
      {Γ(x) = ∀\alist{χ}.τ \\ σ = [\alist{χ}|->\alist{κ}] \\ σ \sim τ }
      {Γ |- x : τσ }
    } \\
    { \inferrule*[Left=LetRef, Right={ρ \textup{fresh}, $ρ\notin τ$}]
      {Γ |-_v t1 : τ' \\ Γ, x : τ'~\refml_ρ |- t2 : τ }
      {Γ |- \letml~x = ref~ t1 ~\inml~ t2 : τ}
    }
  \end{mathpar}
  \caption{A modified and a new rule for aliasing}
  \label{fig:aliasing}
\end{figure}

@@

In the following, we use the metasymbol $v$ to denote well-typed semantic
values, i.e. terms with a typing derivation $Γ|-_v v : τ$.
